---
title: "Co-Kriging_Rincon_Updated"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd('C:/Users/Katie Nelson/Documents')
```

```{r}
my_data=read.csv('JPL_Rincon_March2020FieldCampaign_DATA_EDIT6.csv', header=TRUE);
```


```{r}
library(lattice)
library(gstat)
library(sp)
library(ggplot2)
library(dplyr)
library(caret)
library(AppliedPredictiveModeling)
library(scales)
library(plotly)
library(rgdal)
library(raster)
library(rgeos)
library(dismo)
library(sf)
```

```{r}
xy <- data.frame(my_data)
coordinates(xy) <- c("Long", "Lat")
proj4string(xy) <- CRS("+proj=longlat +datum=WGS84")  ## for example

res <- spTransform(xy, CRS("+proj=utm +zone=16 ellps=WGS84"))
res
#            coordinates ID
# 1 (-48636.65, 1109577)  1
# 2    (213372, 5546301)  2

## For a SpatialPoints object rather than a SpatialPointsDataFrame, just do: 
as(res, "SpatialPoints")
# SpatialPoints:
#              x       y
# [1,] -48636.65 1109577
# [2,] 213372.05 5546301
# Coordinate Reference System (CRS) arguments: +proj=utm +zone=51
# +ellps=WGS84 
```

```{r}
my_data <- as.data.frame(res)
```


```{r}
North_data = slice(my_data, 170:211, 422:793)
NorthLarge_data <- slice(my_data, 422:793)
```


#Create a sampling grid over the study area 
```{r}
NL <- readOGR("LargeNorth_Shapefile.kml")
NL = spTransform(NL, CRS=CRS('+proj=utm +zone=16'))
res <- 30
bb <- bbox(NL)
gt <- GridTopology(cellcentre.offset = bb[,1], cellsize = c(res, res),
                   cells.dim = c(diff(bb[,1]), diff(bb[2,])) / res + 1)
pts <- SpatialPoints(gt, proj4string = CRS(proj4string(NL)))

ov <- sp::over(pts, as(NL, "SpatialPolygons"))
pts_over <- pts[!is.na(ov)]

plot(NL)
points(pts_over)
coordinates(pts_over)
```

```{r}
pt <- data.frame(pts_over)
pt$Lat <- pt$y
pt$Long <- pt$x
```






#Read in NDVI TIFF 
```{r}
str_name<-'20190107_rincon_ndvi.tif' 
ndvi=stack(str_name)
ndvi
```


```{r}
ndvi_crop = crop(ndvi, raster(NL))
```


```{r}
All <- rasterToPoints(ndvi_crop)
```

```{r}
Points_ndvi <- as.data.frame(All)
```


```{r}
Points_ndvi$Lat <- Points_ndvi$y
Points_ndvi$Long <- Points_ndvi$x
```











##North Ridge Co-Kriging 
##Making a data fram with only your x, y, important variables
```{r}
meuse.fg <- NorthLarge_data[seq(1, length(NorthLarge_data$Fg)), c("Long", "Lat", "Fg")]
meuse.fg <- cbind(meuse.fg, ltfg = log10(meuse.fg$Fg))
```

##The same, but for your remotely sensed data 
#multiply by 100 to scale values 
```{r}
meuse.ndvi <- subset(as.data.frame(Points_ndvi), !is.na(X20190107_rincon_ndvi), c("Long", "Lat", "X20190107_rincon_ndvi"))
meuse.ndvi <- cbind(meuse.ndvi, ltndvi = (log10(meuse.ndvi$X20190107_rincon_ndvi)+2),  sltndvi=log10(meuse.ndvi$X20190107_rincon_ndvi)*10)
```


##Telling the program which variables are coordinates for your rs, concentration, and grid data
```{r}
coordinates(meuse.ndvi) <- ~ Lat + Long
coordinates(meuse.fg) <- ~ Lat + Long
coordinates(pt) <- ~ Lat + Long
```

##Calculating your co-kriging parameters step 1
```{r}
g=gstat(NULL, id = "ltfg", form = ltfg ~ 1, data=meuse.fg)
g=gstat(g, id = "ltndvi", form = ltndvi ~ 1, data=meuse.ndvi)
```

##Making sure your values of log transform are positive (you can probably ignore) 
```{r}
meuse.ndvi$ltndvi
```

##Fitting the variogram 
```{r}
v.ltfg <- variogram(ltfg ~ 1, data=meuse.fg)
plot(v.ltfg, pl=T)
```
##fitting the variogram model with lag
```{r}
#lag=c(250, 500, 750, 1000, 1250, 1500, 1750)
lag=c(300, 600, 900, 1200, 1500, 1800)
v.ltfg <- variogram(ltfg ~ 1, data=meuse.fg, width=lag)
plot(v.ltfg, pl=T)
```

##Visual variogram fitting 
```{r}
m.ltfg <- vgm(0.02,"Sph",400,0.05)
plot(v.ltfg, pl=T, model=m.ltfg)
```

##automatic variogram fitting
```{r}
m.ltfg.f <- fit.variogram(v.ltfg, m.ltfg)
plot(v.ltfg, pl=T, model=m.ltfg.f)
```

```{r}
v.ltndvi <- variogram(ltndvi ~ 1, data=meuse.ndvi)
m.ltndvi <- vgm(0.035,"Exp",450,0.001)
plot(v.ltndvi, model=m.ltndvi)
```



##cross variogram for the combo of co2 and rs
```{r}
v.cross <- variogram(g)
plot(v.cross)
```



##applying the variogram model to all the data
```{r}
#g <- gstat(g, id=c("ltfg","sltndvi"), model= m.ltfg.f, fill.all = T)
```

```{r}
g <- gstat(g, id="ltfg", model= m.ltfg.f)
g <- gstat(g, id="ltndvi", model= m.ltndvi)
g <- gstat(g, id=c("ltfg","ltndvi"), model= vgm(0.02,"Sph",400,0.05), fill.all = T)
```

```{r}
plot(v.cross, model=g$model)
```

##fitting the lmc model
```{r}
(g <- fit.lmc(v.cross, g))
plot(variogram(g), model=g$model)
```

##co-kriging between g and your grid file
```{r}
k.c <- predict(g, pt)
str(k.c)
```























#####ADD Code to crop ndvi first!


```{r}
vgm.ndvi <- variogram(X20190107_rincon_ndvi ~ 1, locations = ~Lat + Long, data = Points_ndvi)
vgm.ndvi
plot(vgm.ndvi)
```

```{r}
visual.sph.ndvi <- vgm(0.035, "Sph", 500, 0.001)
plot(vgm.ndvi, visual.sph.ndvi)
```

```{r}
plot(vgm.ndvi) 
auto.sph.ndvi <- fit.variogram(vgm.ndvi, vgm(0.035, "Sph", 500, 0.001))
auto.sph.ndvi
plot(vgm.ndvi, auto.sph.ndvi)
```


Ordinary Kriging
```{r}
# ordinary kriging interpolate
OK_est_ndvi <- krige(NL_est_SGS ~1, locations = ~Lat + Long, data = Points_ndvi, newdata=NorthLarge_data, model=auto.sph.ndvi)
# summary statistics
summary(OK_est_ndvi)
```

```{r}
write.csv(OK_est_ndvi, "ndviPredictOK.csv")
```












