---
title: "Incorporating mature forest as asymptote"
author: "Ana Avila"
date: "July 2024"
format:
  pdf:
    toc: true
---




```{r}
data <- read.csv("data/15y_LULC.csv")
lulc_sum_cols <- grep("lulc_sum_", names(data), value = TRUE)
data$total_LU_sum <- rowSums(data[, lulc_sum_cols], na.rm = TRUE)
data <- subset(data, distance < 5000)
data <- data[!is.na(data$mature_biomass), ]

# Calculate mean and standard deviation per ecoregion
mean_per_ecoregion <- aggregate(mature_biomass ~ ecoreg, data, mean, na.rm = TRUE)
sd_per_ecoregion <- aggregate(mature_biomass ~ ecoreg, data, sd, na.rm = TRUE)

names(mean_per_ecoregion)[2] <- "mean_per_ecoregion"
names(sd_per_ecoregion)[2] <- "sd_per_ecoregion"

data <- merge(data, mean_per_ecoregion, by = "ecoreg", all.x = TRUE)
data <- merge(data, sd_per_ecoregion, by = "ecoreg", all.x = TRUE)

print("secondary forest biomass ~ nearest neighbor mature biomass")
summary(lm(agbd ~ mature_biomass, data = data))$r.squared
print("secondary forest biomass ~ mean mature biomass per ecoregion")
summary(lm(agbd ~ mean_per_ecoregion, data = data))$r.squared
print("nearest neighbor mature biomass ~ distance")
summary(lm(mature_biomass ~ distance, data = data))$r.squared
print("secondary forest biomass ~ age")
summary(lm(agbd ~ age, data = data))$r.squared

plot(data$mature_biomass, data$agbd)
abline(0, 1, col = "red")
pred <- predict(lm(agbd ~ mature_biomass, data = data))
plot(data$agbd, pred)
abline(0, 1, col = "red")
```

With up to 15 years of land use history (up to 20 years of age, including only the youngest forests) the R squared for relationship between secondary forest biomass and nearest neighbor mature biomass is 0.44, about twice that of the average mature biomass per ecoregion. This is after removing all mature forest edge pixels (maintaining only those surrounded by at least a hectare of mature forest) and considering only secondary forests at a hectare of distance or more.

My concern was that the biomass data is averaged out for the edge pixels, but considering this distance there is still a significant relationship between the two.

```{r}
hist(data$agbd, breaks = 100)
hist(data$age)
hist(data$mature_biomass)
hist(data$distance)

data$biomass_distance_interaction <- data$mature_biomass * data$distance
model <- lm(agbd ~ mature_biomass + distance + biomass_distance_interaction, data = data)
summary(model)
```


```{r}
mat_lt_agbd <- data$mature_biomass > data$agbd

# in some cases, the near biomass is actually higher than the agbd.
mat_1sd_avg <- data$mature_biomass < (data$mean_per_ecoregion + data$sd_per_ecoregion) & data$mature_biomass > (data$mean_per_ecoregion - data$sd_per_ecoregion)

combined_condition <- mat_lt_agbd & mat_1sd_avg

percentage_calc <- function(data, condition) {
    num_true <- sum(condition, na.rm = TRUE)
    proportion <- num_true / nrow(data)
    percentage <- proportion * 100
    return(percentage)
}
print("Percentage of pixels where mature forest biomass is higher than secondary forest biomass")
percentage_calc(data, mat_lt_agbd)
print("Percentage of pixels where mature forest biomass is within 1 sd from the ecoregion mean")
percentage_calc(data, mat_1sd_avg)
print("Percentage of pixels where mature forest biomass is higher than secondary forest biomass and mature forest biomass is within 1 sd from the ecoregion mean")
percentage_calc(data, combined_condition)
```

I noticed that for only 68% of the pixels the mature forest biomass is higher than the secondary forest pixels. Even after removing all pixels that pass both conditions (mature forest biomass is higher than secondary forest biomass and mature forest biomass is within 1 sd from the ecoregion mean), the R squared is still 0.46.

```{r echo=TRUE, include=FALSE}
data <- data[combined_condition, ]

summary(lm(agbd ~ mature_biomass, data = data))$r.squared
```
